Arrows for Parallel Computations

### 2017-08-23 paper/content/benchmarks.tex
work in progress on the benchmark chapter. reordering things. SkelRM finished

To showcase that the induced overhead of
PArrows is small, we plot execution time differences between measurements for
PArrows and the corresponding backend in a separate plot
(Figs.~\ref{fig:bench-rm-dist}, \ref{fig:torusBenchmark}). The differences range in
about 0.5~seconds for the execution time of 46~seconds on 256 cores
for distributed \rmtest with PArrows and Eden. The plots show absolute
time differences that are not relativated \wrt the total execution time.

The error bars ends were computed from pointwise maximum of both standard
deviations from both measurements for PArrows and non-PArrows
versions. These are the values provided by the |bench| package that we
used for benchmarking. We call a difference between two versions
significant when the border of the error bar of absolute time
difference is above or below zero. In other words: the time
difference is significant if it is above measurement error.

@@

\olcomment{THE ACTUAL TEXT IS MISSING. What do we see in the plots?
  Why is it good?}
The multicore version of our parallel \rmtest benchmark is in
Figure~\ref{fig:bench-rm-sm}. We have executed the test with 32 and 64
tasks. The plot shows the PArrows-enabled versions with corresponding backends.
The performance of PArrows/Eden~CP in shared memory is slightly better than
for SMP variants such as PArrows/Multicore Haskell and PArrows/|Par|
Monad. The reason for this behavior lies in the backend in our reasoning. One
possible explanation is garbage collection. It is easier to GC
(independently) $n$ heaps than to GC one large. This agrees with known
results\olcomment{WHICH EXACTLY?} that motivate the efforts for
parallel GC in SMP-based Haskells.

To show that PArrows induce very small overhead, we compare the ported
versions of the benchmark to the PArrows-enabled ones with
corresponding backend.\olcomment{THE RESULT?}

In the distributed memory setting, we utilize PArrows/Eden. Eden in
its turn used the MPI backend. We see an almost linear speedup of
\rmtest with 256 tasks and input $2^444497-1$. The sequential run time
was computed as mean of three consecutive executions on a single
core---the single run took two hours 43 minutes. The zero difference  between
PArrows/Eden and Eden almost always lies on the error bar of
the measurement. The only exceptions, where PArrows version was slower
and it was significant, were for 64 and
512 cores, for 0.49 and 0.26~second, respectively. This corresponds to
0.30\% and 0.33\% relative time difference, respectively. The 512 core
version suffers, of course, from the aforementioned hyperthreading
problem. The PArrows-induced overhead was merely 0.56\% for 256
cores. The PArrows version was 0.30\% \emph{faster} for 128 cores.

>>>>

To show that PArrows induce very small overhead in a distributed context as well, we compare the original Eden
versions of the benchmark to its PArrows-enabled counterpart in the \rmtest, \torustest and \jacobitest benchmarks. We plot execution time differences between measurements for
PArrows and the corresponding backend in a separate plot
(Figs.~\ref{fig:bench-rm-dist}, \ref{fig:torusBenchmark}, \ref{fig:bench-jacobi-dist}). As an example, the differences range in
about 0.5~seconds for the execution time of 46~seconds on 256 cores
for distributed \rmtest with PArrows and Eden. For these comparisons, the plots show absolute
time differences that are not relative \wrt the total execution time.
Furthermore, the error bars ends were computed from pointwise maximum of both standard
deviations from both measurements for PArrows and non-PArrows
versions. These are the values provided by the |bench| package that we
used for benchmarking. We call a difference between two versions
significant when the border of the error bar of absolute time
difference is above or below zero. In other words: the time
difference is significant if it is above measurement error.

@@

\olcomment{THE ACTUAL TEXT IS MISSING. What do we see in the plots?
  Why is it good?}
The multicore version of our parallel \rmtest benchmark is in
Figure~\ref{fig:bench-rm-sm}. We executed the test with 32 and 64
tasks. The plot shows the PArrows-enabled versions with corresponding backends.
The performance of PArrows/Eden~CP in shared memory is slightly better than
for SMP variants such as PArrows/Multicore Haskell and PArrows/|Par|
Monad but most of the time the performance is still comparable with the Multicore backend performing slightly worse than the other two in terms of speedup. In particular, the speedups for 16 cores fall behind quite noticeable with 11.69 for 32 tasks as Eden and the |Par| Monad both score 13.05 for 32 tasks. While the Multicore backend scores a bit better in the 64 tasks benchmark, its speedup of 12.70 is still behind the 13.65 and 13.33 of Eden and |Par| Monad, respectively.

Comparing the PArrows version of the \rmtest with the original from Eden with the MPI backend in a distributed memory setting, we see an almost linear speedup of
\rmtest with 256 tasks and input $2^444497-1$ in both versions. The sequential run time
was computed as the mean of three consecutive executions on a single
core---the single run took two hours 43 minutes. The zero difference  between
PArrows/Eden and Eden almost always lies on the error bar of
the measurement. The only exceptions, where PArrows version was slower
and it was significant, were for 64 and
512 cores, for 0.49 and 0.26~second, respectively. This corresponds to
0.30\% and 0.33\% relative time difference, respectively. The 512 core
version suffers, of course, from the aforementioned hyperthreading
problem. The PArrows-induced overhead was merely 0.56\% for 256
cores. The PArrows version was 0.30\% \emph{faster} for 128 cores.
